Actual Causes as Changes of State in Continuous Time

### 2021-01-20 docs/resources/CogSci_paper/paper.tex
pass through expt2

\begin{figure*}[h]
\begin{center}
%\includegraphics[width=7cm]{figures/demo.eps}
\includegraphics{figures/introfig.eps}
\end{center}
\caption{(a) An example of an AND-Gate causal system. In an AND-Gate for an effect to occur it needs the occurence of all the causes it is linked to. Here for the smoke alarm to go off, both the ionization and photoelectric detectors it is connected to have to be activated. In this scenario the detectors activation spreads over the top chain first (ii) and then over the bottom chain leading to the alarm activation (iii).} 
\label{fig:intro}
\end{figure*}

@@

This section uses the smoke detector example in Figure~\ref{fig:intro}a to introduce our theoretical approach to actual causation. Like physical process theory, we focus on processes unfolding in continuous time, but it will be convenient to divide up the dimension of time into a sequence of intervals brief enough that there is at most one event per interval. The effect to be explained is an event that happens within one of these intervals: for example, a smoke alarm's changeof state from inactive to active. Given this setup, our theory can be formulated using a procedure that starts with the effect and steps backwards through time until the actual cause of the effect is identified.

@@

The boundaries of the causal system under consideration will depend on context. For simplicity, our discussion of Figure~\ref{fig:intro}a focused only on events internal to the alarm system, and addressed the question of which detector is best viewed as the main cause of the alarm's activation. Questions about smoke alarms, however, will often be framed with respect to a causal system that includes both the alarm system and its surroundings. Expanding the causal system in this may can allow the alarm's activation to be traced back to an event such as an electrical fault that started a fire which in turn triggered the alarm.

@@

We developed two experiments to test the continuity account of actual causation just presented.  
%The first experiment explores some simple cases directly analogous to the smoke detector example. The second experiment explores cases that elaborate that distinguish between the continuity account and Spellman's probability raising account~\cite{spellman97}.
The cover story for both experiments asked participants to imagine that they worked in a control room and that their job was to monitor networks of particle detectors. Detectors were shown as white squares (as in Figure~\ref{fig:intro} that become active and turn black when they absorb a radioactive particle. Activation is transmitted across links in the network, and a detector activates if all of its input detectors are active. Each network contains a special detector called the Gauge of Critical Moment (GCM, analogous to the smoke alarm in Figure~\ref{fig:intro}a) and participants were told that ``for each activation sequence that you see, your job is to decide what caused the activation of the GCM.''

Both experiments included chains in which the GCM received input from one detector, and dual branch networks such as Figure~\ref{fig:intro}a in which the GCM received input from two detectors. For chains, the continuity account predicts that participants will tend to choose the root cause of the effect (ie the detector that initiates the activation sequence) rather than the immediate cause (the detector that immediately precedes the GCM). For dual branch networks, the continuity account predicts that participants will tend to choose the root cause on the branch whose activation is temporally continuous with the activation of the GCM. We refer to this branch as the continuous branch, and refer to the root cause and the immediate cause on this branch as \ev{R}{C} and \ev{I}{C} respectively, where the subscript denotes ``continuous.'' We refer to the other branch as the ``delayed branch,'' and use \ev{R}{D} and \ev{I}{D} for the root and immediate causes on this branch. The ``delay'' in this naming scheme refers to the delay that occurs between \ev{I}{D} and the activation of the GCM. It is convenient to use the same labels for both detectors and events: for example, \ev{R}{C} will be used to denote both a detector on the continuous branch and the activation of this detector. This flexible use of notation, however, should not be taken to suggest that actual causation involves a relationship between objects instead of a relationship between events.

Our presentation of the experiments focuses on two theoretical accounts: the continuity account and the probability raising account. For completeness, though, we first discuss the counterfactual and physical process accounts. Both accounts can be formulated in different ways --- here we present what we take to be the most natural predictions of the two accounts, and the General Discussion considers ways in which the accounts can be adjusted to make different predictions.

The counterfactual account makes no clear prediction about whether root causes or immediate causes should take priority. For a chain network, if the root cause had not occurred, the effect would not have occurred, but it is equally true that if the immediate cause had not occurred, the effect would not have occurred. If we consider dual branch networks and restrict attention to root causes only, then the account makes no clear preediction about whether \ev{R}{C} should be preferred to \ev{R}{D}.  If \ev{R}{C} had not occurred, then the effect would not have occurred, and likewise for \ev{R}{D}.

The physical process account is similarly inconclusive. Both root and immediate causes are connected by a physical process to the effect, and there seem to be no clear grounds for preferring one to the another. If we focus only on root causes, again there is no clear preference between \ev{R}{C} and \ev{R}{D}. After the activation of \ev{I}{D}, one can think of this detector as continuously sending activation towards the GCM, but this unbroken stream of activation only ``unlocks`` the detector once \ev{I}{C} is also active. It follows that both \ev{R}{C} and \ev{R}{D} are connected to the effect by physical processes.

@@

Each dual branch sequence (short, medium and long) came in four versions. The \emph{state} version showed the delayed branch (including detectors \ev{R}{D} \ev{I}{D}) as active from the very beginning of the sequence. The activation of this branch was therefore presented as having occurred at some indefinite time in the past, resulting in a steady state of activation. The three \emph{event} versions all showed \ev{R}{D} activating 5 seconds into the sequence, and had delays of 2, 4 and 6 seconds between the activation of \ev{I}{D} and \ev{R}{C}.

\textbf{Procedure}.

Participants first read a set of instructions which introduced the task and included examples of a chain and a dual branch network. They then answered three questions about the task and the detectors, and were sent back to read the instructions again if they answered incorrectly. They continued cycling through the instructions and the test questions until they answered all questions correctly.

The 15 activation sequences in the experiment proper were presented in random order. For dual branch sequences, the position of the delayed branch was also randomized (Figure~\ref{fig:intro} shows the delayed branch on the top rather than the bottom). The orientation of the network (GCM on the left or the right) was randomized within participants.  The prompt after each sequence was ``In this sequence what caused the activation of the GCM? Respond by clicking on a detector,'' and participants were required to choose a single detector in response.

\textbf{Results}.

Network orientation (left or right) and position of the delay branch (top or bottom) had no apparent effect, and we therefore collapse across these variables. Figure~\ref{fig:expt1} summarizes the results for short, medium and long dual-branch sequences. The delay between \ev{I}{D} and \ev{R}{C} had no significant effect, and Figure~\ref{fig:expt1} combines results for all three delays.

Although \ev{R}{D}, \ev{I}{D}, \ev{R}{C} and \ev{I}{C} all qualify as causes of the effect, Figure~\ref{fig:expt1} suggests that \ev{R}{C} tends to be singled out as the main cause. This result can be separated into two general conclusions.  First,  participants were more likely to choose root causes than immediate causes. Root and immediate causes were identical for the short sequences and therefore cannot bet distinguished, but the results for medium and long sequences reveal a preference for root causes. STATISTICAL ANALYSIS. The data for chains (not shown in Figure~\ref{fig:expt1}) provide further evidence or a preference for root causes. NUMBERS HERE

@@

Figure \ref{fig:intro} suggests that event and state sequences lead to similar patterns of responses. The continuity account makes the same prediction for both kinds of sequences, but we thought that the event sequences might make participants especially likely to choose \ev{R}{C} over \ev{R}{D}. If this effect exists, our data suggest that it is relatively small.

It is also notable that people's judgments are largely unaffected by the delay between \ev{I}{D} and \ev{R}{C} and the activation length of each sequence (short, medium or long). Both of these variables affect the time that elapses between the root causes and the effect, and our results suggest that people's judgments are not exquisitely sensitive to this kind of variation.

@@

\begin{figure}[ht]
\begin{center}
%\includegraphics[width=8cm]{results_E1}
\includegraphics{figures/expt1.eps}
\end{center}
\caption{Experiment 1. Comparison of the results for different length of activation in AND-Gates. Short: 1 activated detector in each branch. Medium: 3 activated detectors in each branch. Long: 6 activated detectors in each branch. ``State'' responses correspond to the networks where one branch was already activated before participants run the animation. ``Event'' responses correspond to the networks where no detectors were activated before and include results for 2s, 4s and 6s delays between I1 and R2. BUT SEE WARNING ABOVE.} 
\label{fig:2}
\end{figure}

@@

In the previous experiment we have seen that people generally chose R2 (identical to I2 in the short sequences) as the main the cause of the activation of the GCM. These results make sense since R2 and I2 are linked by a continuous sequence of changes. But this set up, where the first element of a chain ``waits'' for the occurrence of the last event of another chain before it can initiate a second sequence, is probably not the most common scenario in real life. For instance it might take more time for one causal process to end so that the causal process that has occurred first is not necessarily the one that ends first. In Experiment 2 we therefore wanted to break the temporal continuity between the root changes and the immediate changes of Experiment 1 such that the continuous sequence of changes be sometimes between R1 and I2 or R2 and I2. We predicted that when there is a continuous sequence of changes between R1 and I2, people would preferably choose R1 as the main cause.

\textbf{Participants}. The experiment was hosted by Google App Engine and 98 participants were recruited via Amazon Mechanical Turk. There were 40 females (average age: 40.4), 57 males (average age: 37.6) and 1 declared ``other''. All the participants were english speakers except for 2 participants.

\textbf{Method}. In Experiment 2 each participant was presented 7 different networks. One stimulus was a chain and all the other six stimuli were AND-Gates. Like in Experiment 1 we tested for the dimension ``Event'' \textit{vs} ``State''. Same as before in condition ``Event'' none of the detectors was initially active and in condition ``State'' one of the two branches was containing detectors that were already active before the animation started. However this time, within the condition ``Event'',  we also tested for the extra dimension ``R1 cont'' \textit{vs} ``R2 cont''.  In condition ``R1 cont'' (represented in Fig.\ref{fig:3}) the activation is such that there is a continuous sequence of changes going from R1 to I2 leading to the activation of G. In condition ``R2 cont'' (not represented here) this is R2 that was continously connected to I2 leading to the activation of G. For the AND-Gates the delay between I1 and I2 in condition ``Event'' was maintained at 11 nodes (i.e.1100 ms) for almost all the networks. There was only one network for which the delay was set at 8 nodes (800ms).

As in Experiment 1 participants were first presented the network in its initial state and were asked to click the ``Run'' button to observe an activation sequence over the network. Participants waited 5000ms to see a first change of state occurring in a detector. Activation delay between any two successive square activations was set on 100ms as well. All the stimuli were presented in a random order and participants were assigned to either group "Left" or "Right" as in Experiment 1, and each AND-Gate was randomly presented with the curved chain (see Fig.\ref{fig:3})) at the top or at the bottom.

@@

\textbf{Results}. First we observed no significant effect (statistics?) of neither the orientation of the networks (left \textit{vs} right) nor the location of the curved chain (top \textit{vs} bottom). On average participants took 8 min to complete the experiment and overall 91\% of the time participants run an animation only once. Second the reversed response pattern for R1 and R2 between ``R1 cont'', ``R2 cont'' and ``State'' stimuli type shown in (Fig.\ref{fig:4}) confirms our hypothesis: when this is the first root change R1 that initiates a continuous sequence of changes until the activation of G, R1 is considered the cause of the activation of G about two times more than R2 (31\% against 15\%) while this is the opposite when this is the second root change that initiates a continuous sequence of changes (18\% against 29\%). Coherently, for ``State'' condition (when this is R2 that initiates the continuous sequence of changes) participants prefer R2 over R1 in a very large extent (47\% against 5\%). As expected the type of network had however no effect on the responses for I1 nor I2 and the percentages of responses for I2 and the root change that initiates the continuous sequence for the Event network (31\% and 34\% for respectively the ``R1 cont'' and ``R2 cont'' networks) were approximately the same.

\begin{figure}[H]
\begin{center}
%\includegraphics[width=8cm]{results_E2}
\includegraphics{figures/expt2.eps}
\end{center}
\caption{Results for AND-Gates where the continuous sequence of changes is either between R1 and I2 or between R2 and I2. ``Event (R1 cont)'': responses for the networks where no detectors were already active before the participant runs the animation and where the continuous sequence of changes is between R1 and I2. ``Event (R2 cont)'': responses for the networks where no detectors were already active and where the continuous sequence is between R2 and I2. ``State (R2cont)'': responses for networks where one branch was already activated before the participant runs the animation and where the continuous sequence of activation is necessarily initiated by R2.} 
\label{fig:4}
\end{figure}

\newpage
\section{General discussion}

>>>>

\begin{figure*}
\begin{center}
%\includegraphics[width=7cm]{figures/demo.eps}
\includegraphics{figures/introfig.eps}
\end{center}
\caption{(a) An example of an AND-Gate causal system. In an AND-Gate for an effect to occur it needs the occurence of all the causes it is linked to. Here for the smoke alarm to go off, both the ionization and photoelectric detectors it is connected to have to be activated. In this scenario the detectors activation spreads over the top chain first (ii) and then over the bottom chain leading to the alarm activation (iii).} 
\label{fig:intro}
\end{figure*}

@@

This section uses the smoke detector example in Figure~\ref{fig:intro}a to introduce our theoretical approach to actual causation. Like physical process theory, we focus on processes unfolding in continuous time, but it will be convenient to divide up the temporal dimension into intervals brief enough that there is at most one event per interval. The effect to be explained is an event that happens within one of these intervals: for example, a smoke alarm's change of state from inactive to active. Given this setup, our theory can be formulated using a procedure that starts with the effect and steps backwards through time until the actual cause of the effect is identified.

@@

The boundaries of the causal system under consideration will depend on context. For simplicity, our discussion of Figure~\ref{fig:intro}a focused only on events internal to the alarm system, and addressed the question of which detector is best viewed as the main cause of the alarm's activation. Questions about smoke alarms, however, will often be framed with respect to a causal system that includes both the alarm system and its surroundings. Expanding the causal system in this way may allow the alarm's activation to be traced back to an event such as an electrical fault that started a fire that generated smoke that triggered the alarm.

@@

We developed two experiments to test the continuity account of actual causation just presented.  
%The first experiment explores some simple cases directly analogous to the smoke detector example. The second experiment explores cases that elaborate that distinguish between the continuity account and Spellman's probability raising account~\cite{spellman97}.
The cover story asked participants to imagine that they worked in a nuclear control room and that their job was to monitor networks of particle detectors. Detectors were shown as white squares (as in Figure~\ref{fig:intro}) that become active and turn black when they absorb a radioactive particle. Activation is transmitted across links in the network, and a detector activates if all of its input detectors are active. Each network contains a special detector called the Gauge of Critical Moment (GCM, analogous to the smoke alarm in Figure~\ref{fig:intro}a) and participants were told that ``for each activation sequence that you see, your job is to decide what caused the activation of the GCM.''

Both experiments included chains in which the GCM received input from one detector, and dual branch networks such as Figure~\ref{fig:intro}a in which the GCM received input from two detectors. For chains, the continuity account predicts that participants will tend to choose the root cause of the effect (ie the detector that initiates the activation sequence) rather than the immediate cause (the detector that immediately precedes the GCM). For dual branch networks, the continuity account predicts that participants will tend to choose the root cause on the branch whose activation is temporally continuous with the activation of the GCM. We refer to this branch as the continuous branch, and refer to the root cause and the immediate cause on this branch as \ev{R}{C} and \ev{I}{C} respectively, where the subscript denotes ``continuous.'' We refer to the other branch as the ``delayed branch,'' and use \ev{R}{D} and \ev{I}{D} for the root and immediate causes on this branch. The ``delay'' in this naming scheme refers to the delay that occurs between \ev{I}{D} and the activation of the GCM. It is convenient to use the same labels for both detectors and events: for example, \ev{R}{C} will be used to denote both a detector on the continuous branch and the activation of this detector. This flexible use of notation, however, does not imply that actual causation involves a relationship between objects rather than events.

Our presentation of the experiments focuses on two theoretical accounts: the continuity account and the probability raising account. For completeness, though, we first discuss the counterfactual and physical process accounts. Both accounts can be formulated in different ways. Here we present what we take to be the default predictions of the two accounts, and the General Discussion considers ways in which the accounts can be adjusted to make different predictions.

The counterfactual account makes no clear prediction about whether root causes or immediate causes should take priority. For a chain network, if the root cause had not occurred, the effect would not have occurred, but if the immediate cause had not occurred, the effect would also not have occurred. If we consider dual branch networks and restrict attention to root causes only, then the account makes no clear prediction about whether \ev{R}{C} should be preferred to \ev{R}{D}.  If \ev{R}{C} had not occurred, then the effect would not have occurred, and likewise for \ev{R}{D}.

The physical process account is similarly inconclusive. Both root and immediate causes are connected by a physical process to the effect, and there seems to be no clear reason for preferring one to the another. If we focus only on root causes, again there is no clear preference between \ev{R}{C} and \ev{R}{D}. After the activation of \ev{I}{D}, one can think of this detector as continuously sending activation towards the GCM which only ``unlocks`` the detector once \ev{I}{C} is also active. It follows that both \ev{R}{C} and \ev{R}{D} are connected to the effect by physical processes.

@@

Each dual branch sequence (short, medium and long) came in four versions. The \emph{state} version showed the delayed branch (including detectors \ev{R}{D} and \ev{I}{D}) as active from the very beginning of the sequence. The activation of this branch was therefore presented as having occurred at some indefinite time in the past, resulting in a steady state of activation. The three \emph{event} versions all showed \ev{R}{D} activating 5 seconds into the sequence, and had delays of 2, 4 and 6 seconds between the activation of \ev{I}{D} and \ev{R}{C}.

\textbf{Procedure}. 
Participants first read instructions which introduced the task and included examples of a chain and a dual branch network. They then answered three questions about the task and the detectors, and were sent back to read the instructions again if they answered incorrectly. They continued cycling through the instructions and the test questions until they answered all questions correctly.

The 15 activation sequences in the experiment proper were presented in random order. For dual branch sequences, the vertical position of the delayed branch was also randomized (Figure~\ref{fig:intro} shows the delayed branch on the top rather than the bottom). The orientation of the network (GCM on the left or the right) was randomized within participants.  The prompt after each sequence was ``In this sequence what caused the activation of the GCM? Respond by clicking on a detector,'' and participants were required to choose a single detector in response.

\textbf{Results}. 
Network orientation (left or right) and position of the delay branch (top or bottom) had no apparent effect, and we therefore collapse across these variables. Figure~\ref{fig:expt1} summarizes the results for short, medium and long dual-branch sequences. The delay between \ev{I}{D} and \ev{R}{C} had no significant effect, and Figure~\ref{fig:expt1} combines results for all three delays.

Although \ev{R}{D}, \ev{I}{D}, \ev{R}{C} and \ev{I}{C} all qualify as causes of the effect, Figure~\ref{fig:expt1} suggests that \ev{R}{C} tends to be singled out as the main cause. This result can be separated into two general conclusions.  First,  participants were more likely to choose root causes than immediate causes. Root and immediate causes were identical for the short sequences and therefore cannot bet distinguished, but the results for medium and long sequences reveal a preference for root causes. STATISTICAL ANALYSIS.

@@

Figure \ref{fig:intro} suggests that event and state sequences lead to similar patterns of responses. The continuity account makes the same prediction for both kinds of sequences, but we anticipated that the event sequences might make participants especially likely to choose \ev{R}{C} over \ev{R}{D}. If this effect exists, our data suggest that it is relatively small.

It is notable that people's judgments are largely unaffected by the delay between \ev{I}{D} and \ev{R}{C} and the activation length of each sequence (short, medium or long). Both of these variables affect the time that elapses between the root causes and the effect, and our results suggest that people's judgments are not exquisitely sensitive to this sort of variation.

Thus far we have focused on dual branch sequences, and the chain sequences are of interest primarily because they provide additional evidence of a preference for root causes.  NUMBERS HERE

@@

\begin{figure}
\begin{center}
%\includegraphics[width=8cm]{results_E1}
\includegraphics{figures/expt1.eps}
\end{center}
\caption{Experiment 1. (a) Order of the four key events for dual branch sequences. The probability raising and continuity accounts both single out \ev{R}{C} as the main cause. (b)  Responses for dual branch sequences. The delayed branch was either inactive (event sequences) or active (state sequences) at the start of the sequence.} 
\label{fig:expt1}
\end{figure}

@@

The results of Experiment 1 are broadly consistent with both the continuity and probability raising accounts, and the goal of Experiment 2 was to distinguish between these accounts. The two make different predictions for dual branch sequences in which the delayed branch activates after activation has already begun along the continuous branch. One such sequence is shown in Figure~\ref{fig:intro}b. 
In cases like this, \ev{R}{D} guarantees the occurrence of the effect, and therefore qualifies as the primary cause according to the probability raising account. The continuity account, however, still treats \ev{R}{C} as the primary cause.

The materials and procedure for Experiment 2 are similar to those for Experiment 1, and we highlight only the few points of difference.

\textbf{Participants}. 100 participants were recruited via Amazon Mechanical Turk and paid XXX for a YYY minute experiment. Two did not submit a valid completion code, leaving 98 sets of responses for analysis.

%The experiment was hosted by Google App Engine and 98 participants were recruited via Amazon Mechanical Turk. There were 40 females (average age: 40.4), 57 males (average age: 37.6) and 1 declared ``other''. All the participants were english speakers except for 2 participants.

\textbf{Design}. 
The experiment included one chain sequence and 6 dual branch sequences. Each dual branch included one straight branch and a longer curved branch, as shown in Figure~\ref{fig:intro}b. The dual branch sequences included the events \ev{R}{C}, \ev{R}{D}, \ev{I}{C}, and \ev{I}{D} in three different orders shown in Figures~\ref{fig:expt2}a and \ref{fig:expt2}b. The sequence in Figure~\ref{fig:intro}b is an instance of the order in Figure~\ref{fig:expt2}a, because the activation on the delayed branch starts (\ev{R}{D}) and finishes (\ev{I}{D}) while activation is propagating along the continuous branch. The 6 dual branch sequences included 2 variants of each of the three orders. In one variant \ev{R}{C} belonged to the curved branch (as in Figure~\ref{fig:intro}b) and in the other \ev{R}{C} belonged to the straight branch.

% XXX: event vs state?

\textbf{Procedure}. 
The position of the curved branch (top or bottom) was randomized. The presentation order of the activation sequences and the orientation (GCM on the left or right) were randomized as for Experiment 1.
%In Experiment 2 each participant was presented 7 different networks. One stimulus was a chain and all the other six stimuli were AND-Gates. Like in Experiment 1 we tested for the dimension ``Event'' \textit{vs} ``State''. Same as before in condition ``Event'' none of the detectors was initially active and in condition ``State'' one of the two branches was containing detectors that were already active before the animation started. However this time, within the condition ``Event'',  we also tested for the extra dimension ``R1 cont'' \textit{vs} ``R2 cont''.  In condition ``R1 cont'' (represented in Fig.\ref{fig:3}) the activation is such that there is a continuous sequence of changes going from R1 to I2 leading to the activation of G. In condition ``R2 cont'' (not represented here) this is R2 that was continously connected to I2 leading to the activation of G. For the AND-Gates the delay between I1 and I2 in condition ``Event'' was maintained at 11 nodes (i.e.1100 ms) for almost all the networks. There was only one network for which the delay was set at 8 nodes (800ms).

%As in Experiment 1 participants were first presented the network in its initial state and were asked to click the ``Run'' button to observe an activation sequence over the network. Participants waited 5000ms to see a first change of state occurring in a detector. Activation delay between any two successive square activations was set on 100ms as well. All the stimuli were presented in a random order and participants were assigned to either group "Left" or "Right" as in Experiment 1, and each AND-Gate was randomly presented with the curved chain (see Fig.\ref{fig:3})) at the top or at the bottom.

@@

\textbf{Results}.  Network orientation (left or right) and position of the curved branch (top or bottom) had no effect and we collapsed across these variables.  Figure~\ref{fig:expt2}c summarizes the results for the three orders listed in Figures \ref{fig:expt2}a and \ref{fig:expt2}b. Consistent with Experiment 1,  \ev{R}{C} is preferred over \ev{R}{D} given the order \{\ev{R}{D},\ev{I}{D},\ev{R}{C},\ev{I}{C}\}, and a similar (but weaker) effect was found for the \{\ev{R}{D},\ev{R}{C},\ev{I}{D},\ev{I}{C}\} sequences. The critical finding is that \ev{R}{C} is also preferred for the \{\ev{R}{C},\ev{R}{D},\ev{I}{D},\ev{I}{C}\} sequences, even though the probability raising account makes the opposite prediction. STATISTICS HERE.

Relative to Experiment 1, the frequency with which \ev{I}{C} is chosen has increased in Experiment 2. This difference may reflect the increased difficulty of Experiment 2. When both branches of a dual branch structure are simultaneously active, keeping track of both root causes and the order in which they occurred is relatively challenging, which may lead some participants to fall back on the simple strategy fo choosing the immediate cause that directly precedes the effect.

%First we observed no significant effect (statistics?) of neither the orientation of the networks (left \textit{vs} right) nor the location of the curved chain (top \textit{vs} bottom). On average participants took 8 min to complete the experiment and overall 91\% of the time participants run an animation only once. Second the reversed response pattern for R1 and R2 between ``R1 cont'', ``R2 cont'' and ``State'' stimuli type shown in (Fig.\ref{fig:4}) confirms our hypothesis: when this is the first root change R1 that initiates a continuous sequence of changes until the activation of G, R1 is considered the cause of the activation of G about two times more than R2 (31\% against 15\%) while this is the opposite when this is the second root change that initiates a continuous sequence of changes (18\% against 29\%). Coherently, for ``State'' condition (when this is R2 that initiates the continuous sequence of changes) participants prefer R2 over R1 in a very large extent (47\% against 5\%). As expected the type of network had however no effect on the responses for I1 nor I2 and the percentages of responses for I2 and the root change that initiates the continuous sequence for the Event network (31\% and 34\% for respectively the ``R1 cont'' and ``R2 cont'' networks) were approximately the same.

\begin{figure}
\begin{center}
%\includegraphics[width=8cm]{results_E2}
\includegraphics{figures/expt2.eps}
\end{center}
\caption{Experiment 2. (a) The probability raising and continuity accounts make different predictions for dual branch sequences in which \ev{R}{C} occurs before \ev{R}{D}. (b) The two accounts agree when \ev{R}{C} occurs after \ev{R}{D}.
(c) Responses for the three possible orders of the key events.} 
\label{fig:expt2}
\end{figure}

\section{General discussion}
